RadLib: a radiative heat transfer model library for CFD

### 2020-12-19 manuscript/manuscript.tex
Edits to model section, all except the RCSLW section; that's next

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}
\usepackage{amsmath}

\usepackage{float}
\restylefloat{table}

@@

% TO DO
% - update README
% - tag software version
% - decide on and generate documentation

@@

@@

The RTE for intensity $I$ in (unit) direction $\mathbf{s}$ for an absorbing, emitting, and non-scattering medium is described by
%
\begin{equation}
	\label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
%
where $I_{\eta}$ is the radiative spectral intensity, 
%$I_{b\eta}$ is the Planck spectral distribution of blackbody intensity, 
$I_{b\eta}$ is spectral blackbody intensity (or Planck function), 
$\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a pathlength along direction $\mathbf{s}$. Radiation property models such as those implemented by RadLib are used to calculate the absorption coefficients $\kappa_{\eta}$. 
%
The gas absorption spectrum is an extremely complex function consisting of millions of spectral lines (peaks) corresponding to quantum mechanical energy levels of the gas molecules.
Direct solution of Eq.~(\ref{e:RTE_general}), involves so-called \emph{line-by-line} calculations that are computationally prohibitive for most practical calculations, and are used primarily for model validation. Global models are an important class of radiation property models that make use of spectrally-integrated radiation properties and are usually versions of the weighted sum of gray gases (WSGG) model \cite{Hottel_1967,Modest_2013}. For instance, integrating Eq.~(\ref{e:RTE_general}) over fixed spectral ranges $\eta_j$ and assuming gray (spectrally uniform) properties in each region $j$, (that is, in each \emph{gray gas} $j$), gives
%
\begin{equation} \label{e:RTE_wsgg}
	\frac{dI_j}{ds} = -\kappa_jI_j + a_j\kappa_jI_b, \hspace{20pt} j=0,1,...,n,
\end{equation}
%
where $a_j$ is a weight factor defined as
%
\begin{align}
    &a_j= \frac{\int_{\eta_j}I_{b,\eta}d\eta}{I_b},\\
    &\sum_j a_j = 1.
\end{align}
%
Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~(\ref{e:RTE_wsgg}) can be computed for each of the $n+1$ gray gases. Normally, the $j=0$ gas is considered as a clear gas and represents \emph{transparent windows} in the spectrum for which $\kappa_0=0$.
The total radiative intensity $I$ is calculated by summing the individual gray gas intensities:
%
\begin{equation}
	I=\int_{\eta=0}^{\infty}I_{\eta}d\eta=\sum_{j=1}^n I_j.
\end{equation}
%
In the above, $I$ is a function of direction $\mathbf{s}$. The heat flux vector $\mathbf{q}$ is found by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere:
%
\begin{equation}
    \mathbf{q} = \int_{4\pi}I(\mathbf{s})\mathbf{s}d\Omega.
\end{equation}
%
The volumetric radiative heat source is then 
%
\begin{equation}
    Q = -\nabla\cdot\mathbf{q}.
\end{equation}
%

@@

RadLib's simplest model uses Planck Mean absorption coefficients calculated from the correlations given on the TNF Workshop site \citep{Smith_2003}. The temperature-dependent correlations are based on the RADCAL model in \citep{Grosshandler_1993}. The TNF radiation model is also documented in \citep{Barlow_2001}.

@@

The basic assumption of weighted sum of gray gases (WSGG) models in general is that the non-gray behavior of gas mixtures, in this case H$_2$O and CO$_2$, can be modeled by a weighted sum of several gray gases and one transparent gas (which represents the spectral windows between absorption bands). 
RadLib implements the WSGG model presented by Bordbar et al. \citep{Bordbar_2014,Bordbar_2020}, which uses correlations based on the HITEMP 2010 database \cite{Rothman_2010}. It uses four gray gases and one transparent gas. Absorption coefficients are calculated by 
\begin{equation}
 \kappa_i=\sum_{k=0}^{4}d_{i,k}M_r^k,
\end{equation}
where $\kappa_i$ is the absorption coefficient for gray gas species $i$, $d_{i,k}$ is a species-specific correlated model coefficient, and $M_r$ is the molar ratio $Y_{\mathrm{H_2O}}/Y_{\mathrm{CO_2}}$. The weight factors are calculated by 
\begin{equation}
 a_i=\sum_{j=0}^{4}b_{i,j}T_r^j,
\end{equation}
where $a_i$ is the weighting factor for species $i$ and $T_r$ is a normalized temperature equal to $T/T_{ref}$ with $T_{ref}=1200$K. The value of $b_{i,j}$ is calculated by 
\begin{equation}
 b_{i,j}=\sum_{k=0}^{4}C_{i,j,k}M_r^k,
\end{equation}
where $C_{i,j,k}$ is another correlated model coefficient and $M_r$ is the molar ratio $Y_{\mathrm{H_2O}}/Y_{\mathrm{CO_2}}$ as above. The model coefficients $d_{i,k}$ and $C_{i,j,k}$ can be found in the literature \citep{Bordbar_2014,Bordbar_2020}.

Previous WSGG versions and other property models often considered weighting factors as a function of temperature only and assumed a fixed molar ratio of H$_2$O and CO$_2$. This version includes dependence on both temperature and an arbitrary molar fraction ratio, increasing its accuracy and flexibility. RadLib uses the updated model coefficients from \citep{Bordbar_2020}, which extends the limited range of H$_2$O-CO$_2$ molar ratios originally presented in \citep{Bordbar_2014}.

@@

The RCSLW model begins with the SLW form of the RTE, given by Equation \ref{e:RTE_SLW}. The gray gas absorption coefficients $\kappa_j$ are calculated calculated as $\kappa_j=N \sqrt{C_{j-1}C_j}$,
where $C_{j-1}$ and $C_j$ are supplemental absorption cross-sections used to discretize the absorption spectrum \cite{Solovjov_2001}. The gray gas weights $a_j$ corresponding to each absorption coefficient are given by $a_j=F(C_j)-F(C_{j-1})$, where $F(C_j)$ is the absorption line blackbody distribution function (ALBDF) for species $j$, calculated from the detailed absorption spectrum of that species. In general SLW methods, the absorption cross-sections are chosen with respect to a thermodynamic reference state, classifying them as reference approaches. The Rank-Correlated SLW (RCSLW) model avoids specifying a reference state by relating thermodynamic states to one another instead. If two arbitrary absorption cross-sections define identical wavenumber intervals, the two cross-sections (neither of which is designated a reference cross-section) can be said to be rank-correlated. The inverse ALBDF, defined such that $C[F(C,\phi_g,T_b),\phi_g,T_b]=C$ and $F[C(F,\phi_g,T_b),\phi_g,T_b]=F$ for a given gas thermodynamic state $\phi_g$ and blackbody temperature $T_b$, is proven to have this property and can also be called the rank-correlated reordered absorption cross-sections \cite{Solovjov_2017}. As a result, the inverse ALBDF can be used to construct cross-section intervals for the RCSLW model that do not rely on an arbitrary reference state or spectrum. ALBDFs for individual gray gases are multiplied together to obtain an overall ABLDF for the gas mixture \cite{Solovjov_2001}. RadLib's RCSLW model uses Method 1.2.2 as defined in and recommended by \cite{Solovjov_2017}.

@@

RadLib can also account for the radiative properties of soot in combustion environments. Soot particles are assumed to be unagglomerated spheres that absorb and emit radiation at all wavelengths such that their optical properties can be determined using the Rayleigh small particle limit. Radiation scattering may be neglected because unagglomerated soot particles in the Rayleigh regime are still small relative to the wavelength of infrared radiation in flames \cite{Brewster_1992,Solovjov_2001}. The spectral soot absorption coefficient is given by 
\begin{equation}
\label{e:soot1}
	\kappa_{\lambda}=C_0 \frac{f_v}{\lambda} , \; \; \; C_0 = \frac{36\pi n k}{(n^2-k^2+2)^2+4n^2k^2},
\end{equation}
where $n$ and $k$ are the real and imaginary parts of the soot complex index of refraction, respectively, and $f_v$ is the soot volume fraction \cite{Solovjov_2001,Modest_2013}. Various values for the refraction index have been proposed and applied to various scenarios \cite{Lee_1981,Stull_1960,Dalzell_1969,Howarth_1966,Chang_1990,Felske_1984,Williams_2007}. RadLib uses the values $n=1.75$ and $k=1.03$ given in \cite{Williams_2007}, which result in a value of $C_0=7.03$, aligning nicely with the value $C_0=7.0$ suggested by \cite{Solovjov_2001}. Due to the simple $1/\lambda$ dependence and the suitably small difference between the mean absorption coefficients calculated for optically thin and optically thick cases, we can use one expression to calculate a spectral mean absorption coefficient for soot that applies to all optical regimes:
\begin{equation}
\label{e:soot2}
	\kappa_{soot}=3.72\frac{f_vC_0T}{C_2},
\end{equation}
where $T$ is the temperature and $C_2=1.4388$ cm K is a Planck function constant \cite{Modest_2013}.

In order to incorporate its radiative effects, soot is modeled as a pseudo-gas of its own and included alongside the existing gray gases. In the Planck Mean model, which models a gas mixture as one gray gas, the soot absorption coefficient calculated from Equation \ref{e:soot2} is simply added to the existing gas absorption coefficient. In the WSGG model, the soot absorption coefficient is similarly calculated from Equation \ref{e:soot2} but is then added to each of the gases separately. Because soot radiates across all wavelengths, its absorption coefficient is added to the values of all four gray gases and the transparent gas that represents the spectral gaps between absorption bands. Unlike the Planck Mean and WSGG models, the RCSLW model requires a spectrally dependent soot model. Instead of finding a mean soot absorption coefficient, we use Equation \ref{e:soot1} to define a soot absorption cross-section and ALBDF as detailed in \cite{Solovjov_2001}. The soot ALBDF can then be multiplied with the other gray gas ALBDFs to find the overall ALBDF for the gas-soot mixture.

@@

Figure \ref{fig:flowchart} illustrates the basic structure use of the RadLib package within a generic example using the provided interfaces. The RadLib library generates absorption coefficients and their weighting factors for use within the appropriate RTE, but does not specify any particular geometry, making it a versatile tool for any simulation that requires radiative heat transfer, regardless of configuration. The interfaces, solver, and examples included with the library serve to illustrate its use and validate the implementation and results.

\begin{figure}
	\begin{center}
		\includegraphics[width=\textwidth]{../figures/radlib_structure.pdf}
	\end{center}
	\caption{Example workflow diagram. Highlighted areas are part of the RadLib library; other areas represent example infrastructure for using the package.}
\label{fig:flowchart}
\end{figure}

>>>>

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.

\usepackage{lineno}
\usepackage{amsmath}
\usepackage{float}
\usepackage[version=4]{mhchem}
\usepackage{booktabs}
\usepackage{siunitx}

\restylefloat{table}

@@

% TODO tag software version

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The RTE for intensity $I$ in (unit) direction $\mathbf{s}$ for an absorbing, emitting, and non-scattering medium is described by
% 
\begin{linenomath}
\begin{equation} \label{e:RTE_general}
    \frac{dI_{\eta}}{ds} = -\kappa_{\eta}I_{\eta} + \kappa_{\eta}I_{b,\eta},
\end{equation}
\end{linenomath}
%
where $I_{\eta}$ is the radiative spectral intensity, 
$I_{b\eta}$ is spectral blackbody intensity (or Planck function), 
$\kappa_{\eta}$ is the spectral absorption coefficient of the medium, and $s$ is a pathlength along direction $\mathbf{s}$. Radiation property models such as those implemented by RadLib are used to calculate the absorption coefficients $\kappa_{\eta}$. 
%
The gas absorption spectrum is an extremely complex function consisting of millions of spectral lines (peaks) corresponding to quantum mechanical energy levels of the gas molecules.
Direct solution of Eq.~(\ref{e:RTE_general}), involves so-called \emph{line-by-line} calculations that are computationally prohibitive for most practical calculations, and are used primarily for model validation. Global models are an important class of radiation property models that make use of spectrally-integrated radiation properties and are usually versions of the weighted sum of gray gases (WSGG) model \cite{Hottel_1967,Modest_2013}. For instance, integrating Eq.~(\ref{e:RTE_general}) over fixed spectral ranges $\eta_j$ and assuming gray (spectrally uniform) properties in each region $j$, (that is, in each \emph{gray gas} $j$), gives
%
\begin{linenomath}
\begin{equation} \label{e:RTE_wsgg}
	\frac{dI_j}{ds} = -\kappa_jI_j + a_j\kappa_jI_b, \hspace{20pt} j=0,1,...,n,
\end{equation}
\end{linenomath}
%
where $a_j$ is a weight factor defined as
%
\begin{linenomath}
\begin{align}
    &a_j= \frac{\int_{\eta_j}I_{b,\eta}d\eta}{I_b},\\
    &\sum_j a_j = 1.
\end{align}
\end{linenomath}
%
Given spatial profiles of $\kappa_j$ and $a_j$, Eq.~(\ref{e:RTE_wsgg}) can be computed for each of the $n+1$ gray gases. Normally, the $j=0$ gas is considered as a clear gas and represents \emph{transparent windows} in the spectrum for which $\kappa_0=0$.
The total radiative intensity $I$ is calculated by summing the individual gray gas intensities:
%
\begin{linenomath}
\begin{equation}
	I=\int_{\eta=0}^{\infty}I_{\eta}d\eta=\sum_{j=1}^n I_j.
\end{equation}
\end{linenomath}
%
In the above, $I$ is a function of direction $\mathbf{s}$. The heat flux vector $\mathbf{q}$ is found by integrating over directional solid angles ($\mathbf{s}\Omega$) on the sphere:
%
\begin{linenomath}
\begin{equation}
    \mathbf{q} = \int_{4\pi}I(\mathbf{s})\mathbf{s}d\Omega.
\end{equation}
\end{linenomath}
%
The volumetric radiative heat source is then 
%
\begin{linenomath}
\begin{equation}
    Q = -\nabla\cdot\mathbf{q}.
\end{equation}
\end{linenomath}
%

@@

RadLib's simplest model uses Planck Mean absorption coefficients calculated from the correlations given on the TNF Workshop site \citep{Smith_2003}. The temperature-dependent correlations are based on the RADCAL model \citep{Grosshandler_1993}. The TNF radiation model is also documented in \citep{Barlow_2001}.

@@

The WSGG model was outlined in Sec.~\ref{s:models}.
RadLib implements the WSGG model presented by Bordbar et al. \citep{Bordbar_2014,Bordbar_2020}, which uses correlations based on the HITEMP 2010 database \cite{Rothman_2010}. It uses four gray gases and one transparent gas, and considers gaseous species \ce{CO2} and \ce{H2O}. This version of the WSGG model has the advantage over some others in that it allows for arbitrary \ce{CO2} and \ce{H2O} compositions, increasing its accuracy and flexibility. Absorption coefficients are calculated by 
%
\begin{linenomath}
\begin{equation}
    \kappa_j=\sum_{k=0}^{4}d_{j,k}M_r^k,
\end{equation}
\end{linenomath}
%
where $\kappa_j$ is the absorption coefficient for gray gas $j$, $d_{j,k}$ is a correlated model coefficient, and $M_r$ is the molar ratio $y_{\mathrm{H_2O}}/y_{\mathrm{CO_2}}$. The weight factors are calculated by 
%
\begin{linenomath}
\begin{equation}
    a_j=\sum_{k=0}^{4}b_{j,k}T_r^k,
\end{equation}
\end{linenomath}
%
where $a_j$ is the weighting factor for gas $j$ and $T_r$ is a scaled temperature equal to $T/T_{ref}$ with $T_{ref}=1200$K. The value of $b_{i,j}$ is calculated by 
%
\begin{linenomath}
\begin{equation}
 b_{j,k}=\sum_{i=0}^{4}C_{j,k,i}M_r^i,
\end{equation}
\end{linenomath}
%
where $C_{j,k,i}$ is another correlated model coefficient. The model coefficients $d_{j,k}$ and $C_{j,k,i}$ are provided by Bordbar et~al. \citep{Bordbar_2014,Bordbar_2020}.

The correlations provided in \cite{Bordbar_2014} are valid for $0.01\le M_r\le4$. In \cite{Bordbar_2020} this range was extended to arbitrary $M_r$ by interpolation of $\kappa_j$ and $a_j$ between the pure component values and the corresponding limit on $M_r$, that is, for \ce{CO2}: interpolation between values at $M_r=0.01$ and pure \ce{CO2}; and for \ce{H2O}: interpolation between values at $M_r=4$ and pure \ce{H2O}.

@@

The RCSLW model begins with the SLW form of the RTE, given by Eq.~(\ref{e:RTE_SLW}). The gray gas absorption coefficients $\kappa_j$ are calculated calculated as $\kappa_j=N \sqrt{C_{j-1}C_j}$,
where $C_{j-1}$ and $C_j$ are supplemental absorption cross-sections used to discretize the absorption spectrum \cite{Solovjov_2001}. The gray gas weights $a_j$ corresponding to each absorption coefficient are given by $a_j=F(C_j)-F(C_{j-1})$, where $F(C_j)$ is the absorption line blackbody distribution function (ALBDF) for species $j$, calculated from the detailed absorption spectrum of that species. In general SLW methods, the absorption cross-sections are chosen with respect to a thermodynamic reference state, classifying them as reference approaches. The Rank-Correlated SLW (RCSLW) model avoids specifying a reference state by relating thermodynamic states to one another instead. If two arbitrary absorption cross-sections define identical wavenumber intervals, the two cross-sections (neither of which is designated a reference cross-section) can be said to be rank-correlated. The inverse ALBDF, defined such that $C[F(C,\phi_g,T_b),\phi_g,T_b]=C$ and $F[C(F,\phi_g,T_b),\phi_g,T_b]=F$ for a given gas thermodynamic state $\phi_g$ and blackbody temperature $T_b$, is proven to have this property and can also be called the rank-correlated reordered absorption cross-sections \cite{Solovjov_2017}. As a result, the inverse ALBDF can be used to construct cross-section intervals for the RCSLW model that do not rely on an arbitrary reference state or spectrum. ALBDFs for individual gray gases are multiplied together to obtain an overall ABLDF for the gas mixture \cite{Solovjov_2001}. RadLib's RCSLW model uses Method 1.2.2 as defined in and recommended by \cite{Solovjov_2017}.

@@

RadLib can also account for the radiative properties of soot in combustion environments. Soot particles are assumed to be unagglomerated spheres that absorb and emit radiation at all wavelengths such that their optical properties can be determined using the Rayleigh small particle limit. Radiation scattering may be neglected because unagglomerated soot particles in the Rayleigh regime are still small relative to the wavelength of infrared radiation in flames \cite{Brewster_1992,Solovjov_2001}. The spectral soot absorption coefficient is given by 
%
\begin{linenomath}
\begin{equation}
\label{e:soot1}
	\kappa_{s,\eta}=C_0 f_v \eta, \; \; \; C_0 = \frac{36\pi n k}{(n^2-k^2+2)^2+4n^2k^2},
\end{equation}
\end{linenomath}
%
where $n$ and $k$ are the real and imaginary parts of the soot complex index of refraction, respectively, and $f_v$ is the soot volume fraction \cite{Solovjov_2001,Modest_2013}. Various values for the refraction index have been proposed and applied \cite{Lee_1981,Stull_1960,Dalzell_1969,Howarth_1966,Chang_1990,Felske_1984,Williams_2007}. 
% INSERT FIGURE SHOWING DIFFERENT MODELS
RadLib uses the values $n=1.75$ and $k=1.03$ given in \cite{Williams_2007}, which result in a value of $C_0=7.03$, which is nearly the same as the value $C_0=7.0$ used in \cite{Solovjov_2001}. 
The Planck mean and Rosseland mean soot absorption coefficients are given in \cite{Modest_2013} as $\kappa_s=3.83f_vC_0T/C_2$ and $\kappa_s=3.60f_vC_0T/C_2$, respectively.
Due to the small difference between the coefficients in these two expressions, a mean value is suggested \cite{Felske_1977,Modest_2013}:
%
\begin{linenomath}
\begin{equation}
\label{e:soot2}
	\kappa_{soot}=3.72\frac{f_vC_0T}{C_2},
\end{equation}
\end{linenomath}
%
where $T$ is the temperature and $C_2=0.014388$ \si{m.K} is a Planck function constant \cite{Modest_2013}.

In the Planck Mean model, which treats a gas mixture as one gray gas, the soot absorption coefficient calculated from Eq.~(\ref{e:soot2}) is added to the gas absorption coefficient. This gives $\kappa_s=1817f_vT$ \si{m^{-1}}.

In the WSGG model, Bordbar suggested incorporating soot as follows \cite{Bordbar_personal}. The soot absorption coefficient is similarly calculated from Eq.~(\ref{e:soot2}) but is then added to each of the gases separately. Because soot radiates across all wavelengths, its absorption coefficient is added to the values of all four gray gases and the clear gas. The gas weights $a_j$ are not changed. This model yeilds the expected behavior for the soot-only and gas-only limits.

The soot model used in the WSGG model can be applied to the RCSLW model as well. However, the formulation of the RCSLW model allows for a spectral soot treatment in the same manner as for the gas species~\cite{Solovjov_2001}. Calculation of the soot ALBDF is facilitated by the simple $\eta$ dependence of $\kappa_{s,\eta}$ in Eq.~(\ref{e:soot1}). The fraction of the black energy at wavenumbers below $\eta$ is computed directly using a series expansion from Chang and Rhee \cite{Chang_1984,Solovjov_2001}. The soot ALBDF is then included in the product of species ALBDFs to obtain the mixture ALBDF.

@@

Figure \ref{fig:flowchart} illustrates the basic structure use of the RadLib package within a generic example using the provided interfaces. The RadLib library generates absorption coefficients and their weighting factors for use within the appropriate RTE, but does not specify any particular geometry, making it a versatile tool for any simulation that requires radiative heat transfer, regardless of configuration. The interfaces, solver, and examples included with the library serve to illustrate its use and validate the implementation and results. 
%
\begin{figure}
	\begin{center}
		\includegraphics[width=\textwidth]{../figures/radlib_structure.pdf}
	\end{center}
	\caption{Example workflow diagram. Highlighted areas are part of the RadLib library; other areas represent example infrastructure for using the package.}
\label{fig:flowchart}
\end{figure}
%
